Missing data: a simple introduction

### 2020-02-25 Draft1.tex
Minor changes

\title{Missing data and techniques for dealing with it}
\author{Marius Dioli}
\date{February 2019}

@@

@@

@@

	\section{Illustrative example}

@@

	\section{The validity of missingness mechanisms}

@@

	\section{What is imputation?}
	Imputation refers to substituting or ''filling in'' missing values with plausible estimates. The main strength of imputation is that it does not change the nature of the subsequent analysis. It makes it much easier for the data processor and analyst to work together since the processor's techniques won't interfere with the analysts'. Another benefit of imputation is that it augments the dataset, meaning that the dataset being analysed is larger than the original dataset, potentially yielding higher performance.

	\section{Techniques of interest}
	To properly evaluate the performance of MI, we should compare it to other similar and commonly used techniques, i.e. techniques that manipulate the dataset by either imputing or removing values. When choosing to apply a method it is important to keep two questions in mind. First, does the method under consideration give us consistent estimators for our model? Second, does the method give us appropriate measures of precision? How the following methods answer these questions is how we will decide the superiority of one method over another.

@@

	Highly robust technique that yields gains even with relatively small portions of missing data.

	Only become feasible in the last 20 years due to computing power required.

	Bayesian approach.

	Between imputation variance

@@

	\subsection{Rubin's rules}

	\subsection{Potential pitfalls}
	When there is missing data in several variables of different types ,for example continuous and binary, and the data is structured, specifying the imputation model can be quite difficult. This is because it has to be a joint model...

@@

@@

	Also check out their motivation for the generating of the data

	\subsubsection{Inference}
	Test statistics
	Different tests

	\subsubsection{Prediction}

	\subsubsection{Classification}

	\subsection{Performance across different data types}
	Continuous
	Categorical (binary, multicategory)
	Time to event
	Timeseries

	\subsection{Sensitivity analysis}

@@

	\bibliography{bib3}	
	\bibliographystyle{apalike}

>>>>

\title{Missing data: a simple introduction}
\author{Marius Dioli}
\date{February 2019}

@@

	To quote MENG: ''imputation is not (merely) a compu-
	tational tool but rather a mode of inference, which al-
	lows hierarchical and sequential input of assessment
	and information'' [Multiple-Imputation Inferences with Uncongenial Sources of Input]

@@

	As detailed in Xie and Meng [DISSECTING MULTIPLE IMPUTATION FROM], much of modern data analysis is done in a multi-phase paradigm, where the distinct and sequential phases data acquisition, data pre-processing, and data analysis, can have different or even contradictory assumptions.

@@

	\subsection{Illustrative example}

@@

	\subsection{The validity of missingness mechanisms}

@@

	\section{Imputation}

	\subsection{What is imputation?}
	Imputation refers to substituting or ''filling in'' missing values with plausible estimates. The main strength of imputation is that it does not change the nature of the subsequent analysis. It makes it much easier for the data processor and analyst to work together since the processor's techniques won't interfere with the analysts'.
	%TODO change the previous sentence to reflect congeniality
	 Another benefit of imputation is that it augments the dataset, meaning that the dataset being analysed is larger than the original dataset, potentially yielding higher performance.

	\subsection{Congeniality: when the imputer's and analysts models differ}
	As mentioned in the introduction, modern data analysis' multiple phases mean that the assumptions underlying each phase and the relationship between the phases is an important object of study.  Meng's [meng 1994] concept of congeniality provides a theoretical framework from which to evaluate these differences.

	Congeniality is an important concept in imputation and has a big impact on the statistical validity of MI. In short, uncongeniality arises when the analysis procedure differs substantially from the imputation model. There are many reasons for such difference, for example, the imputer might wish to include as many variables as possible to achieve an accurate imputation, while the analyst might be interested in only a subset of the covariates and the relationship between these. However, as Meng [1994] states, ''If the imputer's model is reasonably accu-
	rate, then following the multiple-imputation recipe
	prevents the analyst from producing inferences with
	serious nonresponse biases''
	Congeniality [Meng]

	''When the imputation model class and the (embedded) analyst’s model class
	differ, the behavior of Rubin’s rules becomes very complicated, capable of producing inconsistent variance estimators, a matter that has received recurrent
	criticisms'' [Meng 2017]

	\subsection{Techniques of interest}
	To properly evaluate the performance of MI, we should compare it to other similar and commonly used techniques, i.e. techniques that manipulate the dataset by either imputing or removing values. When choosing to apply a method it is important to keep two questions in mind. First, does the method under consideration give us consistent estimators for our model? Second, does the method give us appropriate measures of precision? How the following methods answer these questions is how we will decide the superiority of one method over another.

	There are two other competing methods which we should touch on breifly before continuing. Weighting procedures and likelihood-based methods are ... alternatives to multiple imputation. However, there is reason to prefer MI as the default options over these for a few reasons. First that, given congeniality, MI tends to outperform likelihood-based methods for small sample sizes, while delivering similar performance for large sample sizes \cite{schafer_multiple_2016}.

@@

	Highly robust technique that yields gains even with relatively small portions of missing data. It has only reached desired levels of performance in the last decade due to the computing power required.

	Bayesian approach. Although MI is motivated from the Bayesian perspective, much of today's inference whether in science or industry is dominated by frequentist analyses. This causes some (Reference Fay and stuff from MENG) to question whether inference under MI is valid at all. To address this Meng [1994] developed the concept of congeniality, which we will go into more detail on soon.

	Between imputation variance and within imputation variance

	While MI is motivated from the bayesian perspective, it is most often used with subsequent frequentist inference. Thus ...

@@

	We often want to include as many covariates as possible. X have shown that even mildly correlated auxiliary variables can improve subsequent inference [cite annotadet article]

	\subsection{Rubin's rules and Meng's adaptations}

	Under congeniality, Rubin's rules for calculating standard errors, parameter values, and variance suffice.

	If models are uncongenial, we use Meng's rules.

	\subsection{Potential pitfalls and criticism}
	When there is missing data in several variables of different types ,for example continuous and binary, and the data is structured, specifying the imputation model can be quite difficult. This is because it has to be a joint model...

	. Among all the
	criticisms of multiple imputation, the recent work
	of Fay (1991, 1992) is the most intense, as it is
	directed at the validity of multiple-imputation in-
	ference in practice. Fay provided several examples
	(see also Kott, 1992) to show that the variance esti-
	mator obtained from the repeated-imputation combining rules disagrees asymptotically with the sam-
	pling variance of the repeated-imputation estimator,
	even when the imputation model is correctly speci-
	fied. [MENG]

	%^fixed with congeniality

	Taking all this into account, it is important to see MI in context of the options available to someone working with data. As we have established earlier, simply ignoring missing data can lead  to serious biases, while alternative methods yield similar if not inferior performance in certain cases. As long as a potential user is aware of the potential pitfalls, MI is the superior method for dealing with missing data.

@@

	\subsection{MICE}
	''The
	increased flexibility in modeling these conditional distributions may outweigh
	the lack of clear theoretical justification of the method.'' Rubin

@@

	Also check out their motivation for the generating of the dataa

	In the following examples we will compare our models' performance across different data types. These are continuous, categorical (binary and multicategory), time to event data, and timeseries. The latter two

	\subsubsection{Inference}
	Test statistics
	Different tests
	x
	Congenial vs uncongenial models, see the annotated paper for how they did things with regard to imputation vs analysis.

	\subsubsection{Prediction}
	With large publicly available datasets on the one hand, and laws like GDPR requiring deleting certain kinds of data on the other, multiple imputation seems like a strong candidate for improving the predictive quality of models. Most of the literature focuses on inference, so there is comparatively little comparing the performance of different predictive models. Here we will compare three basic model architectures, multiple linear regression, boosted trees, and a simple neural network, to see how they perform with and without multiple imputation.

	\subsubsection{Classification}
	Like with the prediction case above, classification is an area in which the litterature on missing data is sparse and the same conditions apply. In this context we will compare logistic regression, boosted trees, and neural networks.

	\subsection{Sensitivity analysis and uncongenial models}

@@

	\bibliography{bib2}	
	\bibliographystyle{apalike}
